Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2015-06-11 arxiv_v1/fibonacci.tex
REAL arXiv v1

%------------------------------------------------------------------------------------------------------------%
% Comment fonts
%------------------------------------------------------------------------------------------------------------%

\newcommand{\cggb}[1]{\textcolor{blue}{#1}}
\newcommand{\dude}[1]{\textcolor{red}{#1}}
\newcommand{\stf}[1]{\textcolor{green}{#1}}

@@

We consider a 
2-dimensional system with
point-like anyonic quasiparticle excitations, which exhibit exotic statistics under exchange.
The worldlines of these exchanges form \emph{braids}
in (2+1)-dimensions
which act unitarily on the Hilbert space of degrees of freedom known as the \emph{fusion space}.
To access this space we define
the topological observables of the system: these
are total charge measurements within regions bounded by 
closed loops (Wilson loops).
The possible results of such a charge measurement correspond to different anyon particle types in the model.
These dynamics and observables obey algebraic rules
given by a unitary modular tensor category~\cite{Wang2010b}. 
However, from the perspective of
\emph{topological quantum field theory}
anyons are modeled
as punctures in a surface~\cite{Pfeifer2014}, and it is not
the anyons themselves that braid,
but the manifold in which the anyons reside 
that becomes deformed (twisted) by these braid moves.
This is the perspective we adopt (see \Fref{f:braidloop}).

@@

The defining difference between abelian and nonabelian
anyon theories is that in an abelian theory particle content alone uniquely determines the outcome of total charge measurements.
In contrast, outcomes for nonabelian charge measurements depend on the history of the particles as well as their type.
That is, the fusion space is one-dimensional for abelian anyons, and generally larger for nonabelian anyons.
In particular, we consider a system supporting (nonabelian) Fibonacci
anyon excitations, denoted by $\tau$.
Two such anyons can have total charge
that is either $\tau$ or $\vac$ (vacuum), or any superposition of these, and
so the fusion space in this case is 2-dimensional.
We can represent basis states for this space using diagrams 
of definite total charge for the Wilson loops, 
and arbitrary states as linear combinations of these diagrams:
\begin{align*}
\includegraphics[]{pic-fibonacci.pdf}
\end{align*}
For $n$ anyons of type $\tau$, the 
dimension of this space grows asymptotically as $\varphi^n$, where $\varphi=\frac{1+\sqrt{5}}{2}$ 
is the golden ratio.

@@

Typical noise processes in this kind of system are 
pair-creation, hopping, exchange etc.~of anyons. Such a model was analyzed in detail for Ising anyons in Ref.~\cite{Brell2013}.
It was seen that a simplified noise model consisting of pair-creation events only 
is sufficient to capture the qualitative features of an error-correction simulation, and so for convenience we 
will restrict to pair-creation noise processes in the numerical results presented in this study.

Pair-creation events acting within a single tile do not affect the corresponding observable, and so the only pair-creation events we need explicitly consider act between neighboring pairs of tiles. These are then each associated with an edge of the lattice $\Lambda$, chosen 
uniformly at random. The simulation time, and thus the 
error-correction threshold, will be measured in terms of average number of noise processes 
per edge, as opposed to an iid noise probability per edge. The latter measure is not 
appropriate for error processes in nonabelian anyon models, but the two coincide in the limit 
of low error rates in those cases where they are comparable (see Ref.~\cite{Brell2013}).

@@

In order to perform a logical error on our code, a noise process must have support on a 
a homologically nontrivial region of the lattice.
These correspond to processes in which anyonic charge is transported around a non-trivial loop before 
annihilating to vacuum.

@@

We can guarantee that error-correction will succeed whenever
the action of noise plus decoder 
does not percolate the anyons over the lattice,
but it is possible that percolated events may still result in no error. 
The connection between the percolation threshold and the error-correction threshold 
is not well understood in general~\cite{Hastings2014}, though it is clear 
that our threshold estimate will be a lower bound for 
the true threshold that may be found if all events 
(including those that have percolated) were simulated by an optimal decoder. 
However, calculations of such nontrivial operations for small lattice sizes suggest that almost all such 
processes will indeed result in a logical or leakage error.
As such it is likely that neglecting the full simulation of 
such events does not significantly affect our observed threshold value.

@@

%------------------------------------------------------------------------------------------------------------%
\bibliography{refs2}

>>>>

@@

We consider a 
2-dimensional system with
point-like anyonic quasiparticle excitations, which exhibit exotic statistics under exchange.
The worldlines of these exchanges form \emph{braids}
in (2+1)-dimensions
which act unitarily on the Hilbert space of degrees of freedom known as the \emph{fusion space}.
To describe this space we define
the topological observables of the system: these
are total charge measurements within regions bounded by 
closed loops (Wilson loops).
The possible results of such a charge measurement correspond to different anyon particle types in the model.
These dynamics and observables obey algebraic rules
given by a unitary modular tensor category~\cite{Wang2010b}. 
However, from the perspective of
\emph{topological quantum field theory}
anyons are modeled
as punctures in a surface~\cite{Pfeifer2014}, and it is not
the anyons themselves that braid,
but the manifold in which the anyons reside 
that becomes deformed (twisted) by these braid moves.
This is the perspective we adopt (see \Fref{f:braidloop}).

@@

The defining difference between abelian and nonabelian
anyon theories is that in an abelian theory particle content alone uniquely determines the outcome of total charge measurements.
In contrast, outcomes for nonabelian charge measurements depend on the history of the particles as well as their type.
That is, the fusion space is one-dimensional for abelian anyons, and generally larger for nonabelian anyons.
In particular, we consider a system supporting (nonabelian) Fibonacci
anyon excitations, denoted by $\tau$.
Two such anyons can have total charge
that is either $\tau$ or $\vac$ (vacuum), or any superposition of these, and
so the fusion space in this case is 2-dimensional.
We can represent basis states for this space using diagrams 
of definite total charge for the Wilson loops, 
and arbitrary states as linear combinations of these diagrams:
\begin{align*}
\includegraphics[]{pic-fibonacci.pdf}
\end{align*}
For $n$ anyons of type $\tau$, the 
dimension of the fusion space grows asymptotically as $\varphi^n$, where $\varphi=\frac{1+\sqrt{5}}{2}$ 
is the golden ratio.

@@

Typical thermal noise processes in this kind of system are 
pair-creation, hopping, exchange etc.~of anyons. Such a model was analyzed in detail for Ising anyons in Ref.~\cite{Brell2013}.
It was seen that a simplified noise model consisting of pair-creation events only 
is sufficient to capture the qualitative features of an error-correction simulation, and so for convenience we 
will restrict to pair-creation noise processes in the numerical results presented in this study.

Pair-creation events acting within a single tile do not affect the corresponding observable, and so the only pair-creation events we need explicitly consider act between neighboring pairs of tiles. These are then each associated with an edge of the lattice $\Lambda$, chosen 
uniformly at random to model high-temperature, short timescale thermalizing dynamics. The simulation time, and thus the 
error-correction threshold, will be measured in terms of average number of noise processes 
per edge, as opposed to an iid noise probability per edge. The latter measure is not 
appropriate for error processes in nonabelian anyon models, but the two coincide in the limit 
of low error rates in those cases where they are comparable (see Ref.~\cite{Brell2013}).

@@

In order to perform a logical error on our code, a noise process must have support on a 
a homologically nontrivial region of the manifold.
These correspond to processes in which anyonic charge is transported around a non-trivial loop before 
annihilating to vacuum.

@@

We can guarantee that error-correction will succeed whenever
the action of noise plus decoder 
does not percolate the anyons over the lattice,
but it is possible that percolated events may still result in no error. 
The connection between the percolation threshold and the error-correction threshold 
is not well understood in general~\cite{Hastings2014}, though it is clear 
that our threshold estimate will be a lower bound for 
the true threshold that may be found if all events 
(including those that have percolated) were simulated. 
However, calculations of homologically nontrivial operations for small lattice sizes suggest that almost all such 
processes will indeed result in a logical or leakage error.
As such it is likely that neglecting the full simulation of 
these events does not significantly affect our observed threshold value.

@@

%------------------------------------------------------------------------------------------------------------%
\bibliography{fibonacci}
